Le titre

### 2022-10-10 Manuscrit Thèse/chapter 1/Chapter 1.tex
Le gros benj
gbrei

\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme

@@

\subsubsection{CVD growth}

@@

\subsection{Experimental setup}
\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{Figures/shema_exp}
\caption{Confocal microscopy setup detailed in main text}
\label{setup}
\end{figure}
The three most common experimental degrees of freedom to consider for NV experiments are: the optical illumination, the microwave field and the external magnetic field. Other parameters such as the crystal temperature, the electric field or the crystal stress can also be controlled for specific applications.

Fig. \ref{setup} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms. Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.

@@

\subsection{NV spin $T_2$}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/Echo}
\caption{Spin-echo protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
\label{Echo}
\end{figure}
\subsubsection{Measurement protocol}
As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process which is the spin-echo or Hahn-echo sequence.

@@

\subsection{Cross-relaxations with NV centers}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/Shema_CR}
\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
\label{CR_shema}
\end{figure}

>>>>

@@

\subsubsection{CVD growth}
\label{fab HPHT}

@@

\subsection{Experimental setup}
\label{Sec setup cahp 1}
\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{Figures/shema_exp}
\caption{Confocal microscopy setup detailed in main text}
\label{setup ch1}
\end{figure}
The three most common experimental degrees of freedom to consider for NV experiments are: the optical illumination, the microwave field and the external magnetic field. Other parameters such as the crystal temperature, the electric field or the crystal stress can also be controlled for specific applications.

Fig. \ref{setup ch1} shows a multi-purpose experimental setup which offers a control over these three degree of freedoms. Unless specified otherwise, this is the experimental setup that was used to perform all the measurements shown in this manuscript.

@@

\subsection{NV spin $T_2$}
\label{sec T2 echo}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/Echo}
\caption{Spin-echo protocol. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink, fitted with an exponential decay of time $3.9\ \rm{\mu s}$.} 
\label{Echo}
\end{figure}
\subsubsection{Measurement protocol}
As precised earlier, the exact $T_2$ definition depends on the rephasing scheme used. We will focus here on the simplest known rephasing process which is the spin-echo or Hahn-echo sequence.

@@

\subsection{Cross-relaxations with NV centers}
\label{Sec_CR}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/Shema_CR}
\caption{Illustration of cross-relaxations between an NV center and an unpolarized spin. Red dots represent the population in each states, and the arrows the various population transfer mechanisms}
\label{CR_shema}
\end{figure}

### 2022-10-10 Manuscrit Thèse/chapter 2/Chapter 2.tex
gbrei

ODMR can also be significantly more sensitive than EPR: single NV centers are routinely observed in ODMR while the the current EPR noise floor for electronic spins in diamond is typically of $\sim 10^{11}$ \citep{mitchell2013x}. This difference in sensitivity comes in part from the better spin polarization \footnote{At room temperature, the thermal polarization of a transition at 9500 MHz is $\sim 0.002$ while the NV center optical polarization can reach 0.8.}, and from the more efficient photon detection in the optical range than in the microwave range.

EPR however remains a far more versatile tool as only very few spin defects have the properties required for ODMR. The key idea of this chapter is to allow the optical detection of spin defects that do not show ODMR, by making them resonant with NV centers and observing NV cross-relaxation.

@@

 Fig. \ref{T1 vs PL} shows a simulation of NV centers PL when the spin decay rate $\Gamma_1$ is modified. In this fictional scenario, we assume that $\Gamma_1$ resonantly increases for a given value of the magnetic field (we will see later in this manuscript situations matching this scenario). We can see that the PL decreases as an answer to the $\Gamma_1$ increase. The exact value of the PL drop ($\sim 2\%$ in the crase presented here) depends on the optical polarization rate $\Gamma_{\rm las}$, but this value generally does exceed a few \%, even for important $T_1$ modifications.

It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the PL case, the laser intensity needs to be adjusted so that $\Gamma_{\rm las} \sim \Gamma_1$ in order to maximize the PL contrast, while the single$-\tau$ method requires a waiting time $\tau \sim T_1$ between each measurement, which result in a similar PL count and shot noise limit in both case.

On a technical level, the PL-based method is easier to implement since it does not require the use of a pulsed laser, and uses an overall smaller laser power, however it is more sensitive to drifts and changes in the optical setup. Most of the relaxometry measurement in this manuscript were performed using a PL-based detection.

@@

This section focuses on particular field orientation with respect to the diamond crystalline axes, and details the advantages and disadvantages of the two main choices that are the [100] and [111] direction in order to detect four particular dark spins: P1 centers, Vh$^-$, War1 and the $^{13}C-NV$ complex\footnote{The $^{13}C-NV$ complex is normally not dark, as discussed in its own section.}.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/121_vs_22_NV}
\caption{Simulated transition frequencies of the $\ket{0} \to {-1}$ and $\ket{0} \to {+1}$ transitions for the 4 classes of NV centers as a function of the magnetic field. a) When the magnetic field is along the diamond [100] crystalline axis, b) along the [111] axis. A representation of the 4 classes of NV centers, labeled 1 to 4, and the orientation of the magnetic field in both cases is put on top}
\label{121 vs 22 NV}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/121_vs_22_P1}
\caption{Simulation of the transition frequencies of NV$^-$ centers and P1 centers (without the hyperfine interactions) for a magnetic field a) along the [100] axis and b) along the [111] axis}
\label{121 vs 22 P1}
\end{figure}

>>>>

ODMR can also be significantly more sensitive than EPR: single NV centers are routinely observed in ODMR while the the current EPR noise floor for electronic spins in diamond is typically of $\sim 10^{11}$ spins \citep{mitchell2013x}. This difference in sensitivity comes in part from the better spin polarization \footnote{At room temperature, the thermal polarization of a spin with a transition at 9500 MHz is $\sim 0.002$ while the NV center optical polarization can reach 0.8.}, and in part from the more efficient photon detection in the optical range than in the microwave range.

The key idea of this chapter is to allow the optical detection of spin defects that normally do not show ODMR thanks to cross-relaxation with NV centers.

@@

 Fig. \ref{T1 vs PL} shows a simulation of NV centers PL when the spin decay rate $\Gamma_1$ is modified. In this fictional scenario, we assume that $\Gamma_1$ resonantly increases for a given value of the magnetic field (we will see later in this manuscript physical examples corresponding to this scenario). We can see that the PL decreases as an answer to the $\Gamma_1$ increase. The exact value of the PL drop ($\sim 2\%$ in the crase presented here) depends on the optical polarization rate $\Gamma_{\rm las}$ as well as the modification of the spin lifetime.

It is argued in \citep{finco2021imaging} that both methods result in similar signal to noise ratio: in the PL case, the laser intensity needs to be adjusted so that $\Gamma_{\rm las} \sim \Gamma_1$ in order to maximize the PL contrast, while the single$-\tau$ method requires a waiting time $\tau \sim T_1$ between each measurement, which results in a similar PL count and shot noise limit in both case.

On a technical level, the PL-based method is easier to implement since it does not require the use of a pulsed laser, and because it requires a smaller laser power to work optimally. It is however more sensitive to drifts and changes in the optical setup. Most of the relaxometry measurement in this manuscript were performed using a PL-based detection.

@@

This section focuses on particular field orientation with respect to the diamond crystalline axes, and details the advantages and disadvantages of the two main choices that are the [100] and [111] direction in order to detect four particular dark spins: P1 centers, Vh$^-$, War1 and the $^{13}$C$-$NV complex\footnote{The $^{13}$C$-$NV complex is normally not dark, as will be discussed in its own section.}.

@@

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{Figures/121_vs_22_NV}
\caption{Simulated transition frequencies of the $\ket{0} \to {-1}$ and $\ket{0} \to {+1}$ transitions for the 4 classes of NV centers as a function of the magnetic field. a) When the magnetic field is along the diamond [100] crystalline axis, b) along the [111] axis. A representation of the 4 classes of NV centers, labeled 1 to 4, and the orientation of the magnetic field in both cases is put on top}
\label{121 vs 22 NV}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{Figures/121_vs_22_P1}
\caption{Simulation of the transition frequencies of NV$^-$ centers and P1 centers (without the hyperfine interactions) for a magnetic field a) along the [100] axis and b) along the [111] axis}
\label{121 vs 22 P1}
\end{figure}

### 2022-10-10 Manuscrit Thèse/chapter 3/Appendix_eta.tex
gbrei

\begin{equation}
\eta^2=\frac{1}{3} (\abs{g}^2+\abs{h}^2)  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
\end{equation}

\section{$\bar \eta$}

\printbibliography
\end{document}

>>>>

where $\eta$ is defined as:

\begin{equation}
\label{eq. eta debut annexe}
\eta^2=\frac{1}{3} \left(\frac{\mel{i}{\mathcal{H_{\rm dd}}}{f}}{\frac{J_0}{r^3}}\right)^2 \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2},
\end{equation}

and where $\mathcal{H_{\rm dd}}$ is the dipole-dipole Hamiltonian and $\ket{i}$ and $\ket{f}$ the initial and final two-qubits states of the flip-flop or double flip process.

\section{$\bar \eta$ in the magnetic basis $\{ \ket{0}, \ket{+1}, \ket{-1}\}$}

\begin{figure}[h]
\centering
\includegraphics[width=.7\textwidth]{Figures/Shema_annexe}
\caption{Schematics of the ``normal" NV center NV$_s$ and the fluctuator NV$_f$ in and their respective basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$), as well as the relative position $\mathbf{r}(r,\theta,\phi)$ between the two spins.}
\label{shema spins annexe}
\end{figure}

The computation of $\bar \eta$ when the single spin Hamiltonian of each spin is diagonal in the magnetic basis $\{ \ket{0}, \ket{+1}, \ket{-1}\}$ was treated in \citep{choi2017depolarization}. We will summarize their results here and consider a few different geometries than the ones presented in the article.

Fig. \ref{shema spins annexe} represents the two spins and the three relevant angles in the problem $\theta,\phi$ and $\psi$. We label with $s$ the properties associated with the ``normal " NV center and $f$ those associated with the fluctuator. For instance the two-qubits state $\ket{m_s=0,m_f=+1}\equiv\ket{0,+1}$ corresponds to the convolution of the single-spin states $\ket{m=0}$ for the NV and $\ket{m=+1}$ for the fluctuator.

The two Cartesian basis ($\hat x_s,\hat y_s,\hat z_s$) and ($\hat x_f,\hat y_f,\hat z_f$) were chosen so that the spin vector Hamiltonian of each spin could be written:
\begin{equation}
\mathbf{S}_i=S_x \hat x_i + S_y \hat{y}_i + S_z \hat{z}_i,
\end{equation}
where :
\begin{align*}
S_x&=\begin{pmatrix}
0&1/\sqrt{2}&0 \\
1/\sqrt{2}&0&1/\sqrt{2} \\
0&1/\sqrt{2}&0
\end{pmatrix} \\
S_y&=\begin{pmatrix}
0&i/\sqrt{2}&0 \\
-i/\sqrt{2}&0&i/\sqrt{2} \\
0&-i/\sqrt{2}&0
\end{pmatrix} \\
S_z&=\begin{pmatrix}
-1&0&0 \\
0&0&0 \\
0&0&+1
\end{pmatrix}.
\end{align*}

The relative position between the two spins is noted with the vector $\mathbf{r}=r\hat{u}$. The vector's spherical coordinate are given in the Cartesian basis of the NV center ($\hat x_s,\hat y_s,\hat z_s$).

The angle $\psi$ between the two spins $z$ axes is defined by the crystal lattice and can take three values within [0,$\pi$] which are: 0 (same class), $\cos^{-1}(1/3) \approx 70.5\ ^\circ$ and $\cos^{-1}(-1/3) \approx 109.5\ ^\circ$. Two distinct classes can either have a relative angle $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$ depending on the external magnetic field: we choose by definition that $\mathbf{B}\cdot\hat{z}_i >0$, which imposes the orientation of $z_i$ and therefore the angle between $\hat z_s$ and $\hat z_f$. This point will be further developed when discussing the $\{100\}$ and $\{110\}$ type resonances.

The direction $\hat x_s=\hat x_f$ is chosen arbitrarily to be : $$\hat x=\frac{\hat z_s \times \hat z_f}{\norm{\hat z_s \times \hat z_f}}.$$
The arbitrary choice of the $\hat x$ direction is justified by the symmetry of the problem in the ($\hat x, \hat y$) plane. We chose here to neglect the symmetry-breaking part of the transverse magnetic field which is justified by the fact that we do not take into consideration the mixing of the eigenstates caused by the transverse field.

Following the notation introduced in \citep{choi2017depolarization}, we can rewrite the dipole-dipole Hamiltonian as:
\begin{equation}
\mathcal{H}_{\rm dd} \approx -\frac{J_0}{r^3}\left[(g+ih)(\ket{0,+1}\bra{+1,0}+\ket{0,-1}\bra{-1,0}) + h.c. + q S_z^sS_z^f \right],
\end{equation}
with 
\begin{align}
g&=\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{x}_f) - (\hat{x}_s \cdot \hat{x}_f) + 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{y}_s \cdot \hat{y}_f) \right] \\
h&=\frac{1}{2}\left[3(\hat{u}\cdot\hat{x}_s)(\hat{u}\cdot\hat{y}_f) - (\hat{x}_s \cdot \hat{y}_f) - 3(\hat{u}\cdot\hat{y}_s)(\hat{u}\cdot\hat{x}_f) + (\hat{y}_s \cdot \hat{x}_f) \right] \\
q&= 3(\hat{u}\cdot \hat{z}_s)(\hat{u}\cdot \hat{z}_f) - (\hat{z}_s \cdot \hat{z}_f).
\end{align}

We should note that the double flips have been omitted because we consider here the case $B\neq0$ for which the $\ket{+1}$ and $\ket{-1}$ states of each spin are far out of resonance.

Eq. \ref{eq. eta debut annexe} can then be written:
\begin{equation}
\eta^2=\frac{1}{3} (\abs{g}^2 + \abs{h}^2) \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
\end{equation}

In order to compute $\bar \eta$, we first decompose $\eta$ as a product of $R$ and $G$ as defined in [REF]:
\begin{equation}
\eta^2=G^2(\theta, \phi, \psi) R^2(\omega_f,\omega_{NV})
\end{equation}
with
\begin{align*}
G^2(\theta, \phi, \psi)&=\frac{1}{3}\left(\abs{g}^2+\abs{h}^2 \right),  \\ 
R^2(\omega_f,\omega_{NV}) &= \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}.
\end{align*}

$R$ and $G$ can be averaged separately as they do not depend on the same variables, which means that:
\begin{equation}
\label{eq. eta bar r bar g bar}
\bar \eta = \bar R \cdot \bar G.
\end{equation}

The computation of $\bar{R}$ has been discussed in [REF], we will focus here on $\bar{G}$ defined as:

\begin{equation}
\label{eq. eta base mag}
\bar{G}= \iint \abs{G} \frac{\rm d\theta \cos \theta \rm d\phi}{4 \pi}.
\end{equation}

Eq. \ref{eq. eta base mag} can be solved analytically for the case $\psi=0$ and numerically for $\psi=70.5 \ ^\circ$ or $\psi=109.5 \ ^\circ$. The three values are reported in Table \ref{table G mag}. The values found for $\psi=0$ and $\psi=70.5 \ ^\circ$ are the same that were found by the authors of \citep{choi2017depolarization} \footnote{They differ by a factor $\sqrt{1/3}$ because of the definition of $G$.}.

\begin{table}[htbp]
\centering
\caption{Computation of $\bar G$ in the magnetic basis}
 \label{table G mag}
\begin{tabular}{c|c|c}
\toprule
$\psi=0$ & $\psi=70.5 \ ^\circ$ & $\psi=109.5 \ ^\circ$ \\

$\frac{2}{9}=0.222$ & 0.3757 & 0.4808 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Interclass resonance and magnetic field orientation}

\begin{figure}[h]
\centering
\includegraphics[width=.8\textwidth]{Figures/110_vs_100}
\caption{(a) Illustration of the co-resonance condition between a NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis. The $\hat z$ direction is defined by the condition $\mathbf{B}\cdot\hat{z} >0$. (b) 8 individual $T_1$ measurements on sample ADM-15-3 on a two-class resonance for 8 different values of the magnetic field (4 $\mathbf{B} \in \{110\}$ and 4 $\mathbf{B} \in \{100\}$).}
\label{110 vs 100 annexe}
\end{figure}

As discussed in sec [REF], there are 4 possibles magnetic field orientations for which at least two NV classes are co-resonant: $\mathbf{B} \in \{110\}$ (two-class resonance), $\mathbf{B} \in \{100\}$ (2$\times$two-class resonance), $\mathbf{B} \parallel \langle 111 \rangle$ (three-class resonance) and $\mathbf{B} \parallel \langle 100\rangle$ (four-class resonance).

Fig. \ref{110 vs 100 annexe}-a) illustrates the difference in the angle $\psi$ for the case $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$. We consider here an NV center parallel to the [111] axis and a fluctuator parallel to the [$\bar 1$11] axis. Without external fields, the $\hat z$ direction of the NV center could be either [111] or [$\bar 1 \bar 1 \bar 1$], however the condition $\mathbf{B}\cdot\hat{z} >0$ imposes in the left case $\hat z_s=[111]$ and in the right cases $\hat z_f=[\bar 1 \bar 1 \bar 1]$.

For the NV center and the fluctuator to be resonant, the magnetic field needs to have the same projection on the $\hat z_s$ and $\hat z_f$ axes, which in this case means either $\mathbf{B} \in (100)$ or $\mathbf{B} \in (011)$. The angle $\psi$ between the $\hat z_s$ and $\hat z_f$ axes however is different in both cases. The same can be generalized for every $\mathbf{B} \in \{110\}$ and $\mathbf{B} \in \{100\}$ type resonance: the angle $\psi \in [0,\pi]$ between two resonant classes is equal to $70.5 \ ^\circ$ for a $\mathbf{B} \in \{100\}$ type resonance, and $109.5 \ ^\circ$ for a $\mathbf{B} \in \{110\}$ type resonance.

Because $\bar G$ is greater for $\psi=109.5 \ ^\circ$ than it is for $\psi=70.5 \ ^\circ$, then following eq. \ref{eq. eta bar r bar g bar}, we should expect a faster depolarization rate when $\mathbf{B} \in \{110\}$ than when $\mathbf{B} \in \{100\}$. Fig. \ref{110 vs 100 annexe}-b) Shows 8 different $T_1$ measurement performed on the same sample for 8 different magnetic fields, four of which were in the $\mathbf{B} \in \{110\}$ scenario and four in the $\mathbf{B} \in \{100\}$ scenario. We can clearly see a slower relaxation rate for all four $\mathbf{B} \in \{100\}$ measurements compared to all four $\mathbf{B} \in \{110\}$, which agrees with the predictions of the model.

For a perfect resonance matching (i.e. no detuning between the central frequencies of the different resonant classes), we can assume that $\bar R$ is a constant ($\bar R \equiv \bar R^0 \approx \frac{2 \gamma_f}{2 \gamma_f + \Gamma_f + \Gamma_{NV}}$ according to eq. [REF]). We then have:

\begin{align}
\bar \eta(\rm{1\ class})&=\bar G(\psi=0) \bar R_0 \\[1em]
\frac{\bar \eta(\mathbf{B} \in \{110\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
\frac{\bar \eta(\mathbf{B} \in \{100\})}{\bar \eta(\rm{1\ class})}&= \frac{\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
\frac{\bar \eta(\mathbf{B} \parallel \langle 111 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=109.5 \ ^\circ)+ \bar G(\psi=0)}{\bar G(\psi=0)}\\[1em]
\frac{\bar \eta(\mathbf{B} \parallel \langle 100 \rangle)}{\bar \eta(\rm{1\ class})}&= \frac{2\bar G(\psi=70.5 \ ^\circ)+ \bar G(\psi=109.5 \ ^\circ) + \bar G(\psi=0)}{\bar G(\psi=0)}
\end{align}

Finally, the values in Table [REF] are computed thanks to the relation: \begin{equation}
\frac{\Gamma_1}{\Gamma_0}= \left(\frac{\bar \eta}{\bar \eta(\rm{1\ class})} \right)^2.
\end{equation}

Note that we only give here the relaxation rates for a single class. To estimate the total PL, one would need to compute the relaxation rate of all 4 classes independently, for instance in a $\mathbf{B} \in \{110\}$ scenario, two classes have a relaxation rate $\Gamma(\mathbf{B} \in \{110\})$ and two classes have a relaxation rate $\Gamma_0$.
\printbibliography
\end{document}
